---
title: "Question 2"
author: "Your Name"
date: today
format: 
  html:
    theme: cosmo
    code-fold: true
    toc: true
    toc-location: left
    number-sections: true
execute:
  echo: true
  warning: false
---

## Introduction

In this analysis, we will... [Describe the purpose of your analysis and what questions you aim to answer. Provide context about the data and methods you'll be using.]

```{r}
#| label: setup
#| include: false

# Load required packages
library(tidyverse)
library(ggplot2)
library(dplyr)

# Set default theme for plots
theme_set(theme_minimal())
```

## Data Preparation

```{r}
#| label: data-prep
#| message: false

# Load and prepare your data here
# Example:
# data <- read_csv("your_data.csv")
# Clean and transform your data as needed
```

## Exploratory Data Analysis

```{r}
#| label: eda
#| message: false
#| fig-cap: "Description of your visualization"

# Create exploratory visualizations
# Example:
# ggplot(data, aes(x = variable1, y = variable2)) +
#   geom_point() +
#   labs(title = "Title",
#        x = "X-axis label",
#        y = "Y-axis label")
```

## Statistical Analysis

```{r}
#| label: analysis
#| message: false

# Perform your statistical analysis here
# Example:
# model <- lm(y ~ x, data = your_data)
# summary(model)
```

## Results

[Describe your results here. Include:
- Key findings from your exploratory data analysis
- Results from statistical tests
- Interpretation of your findings
- Any interesting patterns or relationships discovered]

## Discussion

[Discuss the implications of your results:
- What do your findings mean in the context of your research question?
- How do they compare to existing literature or expectations?
- What are the limitations of your analysis?]

## Conclusion

[Summarize your main findings and their significance. Include:
- Brief recap of key results
- Broader implications
- Potential future directions]

## References

[Include any references cited in your analysis]